---
title: "process_qualtrics"
output: html_document
date: "2023-11-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup Data Using Qualtrics API

* Load data directly from Qualtrics
* Can store API Key and credentials in `.renviron`

```{r}
library(tidyverse)
library(qualtRics)
library(magrittr)
library(RColorBrewer)

source("../plotting_functions.R")

# datacenterid <- ""
# url <- str_glue("https://{datacenterid}.qualtrics.com")
# api_key <- ""
# 
# qualtrics_api_credentials(api_key = api_key,
#                           base_url = url,
#                           install = TRUE,
#                           overwrite=T)

readRenviron("~/.Renviron")
# load survey data
surveys <- all_surveys()
# select political candidates survey ID
pc_id <- surveys[surveys["name"] == "Political Candidates", ][["id"]]

pc_survey <- fetch_survey(
  surveyID = pc_id,
  verbose = TRUE
)
```

## Survey Validation

```{r}
pc_survey %>%
  distinct(Status)

# validate survey logic on test data
pc_survey %>%
  filter(Status == "Survey Test")

# should be missing if do not consent
pc_survey %>%
  filter(Status == "Survey Test", Consent == "I do not consent to participate") %>%
  mutate(Q2 = as.character(Q2)) %>% 
  distinct(`Q2`)  %>% 
  is.na() %>% 
  stopifnot()

# note: question was different prior to 11/06
# check that if failing prescreen question about location in the US
# this should be missing
pc_survey %>%
  filter(
    Status == "Survey Test", PreScreen_Q1 != "Yes",
    StartDate >= lubridate::ymd("2023-11-06")
  ) %>%
  mutate(Q1 = as.character(Q1)) %>% 
  distinct(Q1) %>% 
  is.na() %>% 
  stopifnot()
```

# Examine the test data

* Check if `rnum` and `pi_i` values are working properly

```{r}
pc_survey_test <- pc_survey %>%
  filter(Status == "Survey Test") %>%
  # create 'profile' variable
  mutate(profile = case_when(
    rnum <= pi1 ~ 1,
    rnum > pi1 & rnum <= pi2 ~ 2,
    rnum > pi2 & rnum <= pi3 ~ 3,
    rnum > pi3 & rnum <= pi4 ~ 4,
    rnum > pi4 & rnum <= pi5 ~ 5,
    rnum > pi5 & rnum <= pi6 ~ 6,
    rnum > pi6 & rnum <= pi7 ~ 7,
    rnum > pi7 ~ 8
  ))


# check distributoin
# this should be evenly distributed
# each should be about 1/8 of the data
nrow(pc_survey_test)/8

pc_survey_test %>%
  ggplot(aes(profile)) +
  geom_histogram() +
  theme_classic()
```
```{r}
# create numeric versions of the profiles
# should be =1 if selecting the younger candidate
# this depends on whether the younger candidate
# is candidate 1 or candidate 2
pc_survey_recode <- pc_survey_test %>%
  mutate(across(str_c("Q", 1:8), ~ case_when(
    . == "Candidate 1" & rnum_age <= 0.5 ~ 1,
    . == "Candidate 2" & rnum_age > 0.5 ~ 1,
    TRUE ~ 0
  ))) %>%
  mutate(total_Q = select(., str_c("Q", 1:8)) %>%
    rowSums())

# check row total is not more than 1
# 0 if selected the older candidate
pc_survey_recode %>%
  select(str_c("Q", 1:8)) %>%
  rowSums() %>% 
  is_weakly_less_than(1) %>% 
  all() %>% 
  stopifnot()
```

## Create fake data

```{r, setup fake data}
# create fake data
# batch size is the size of the batches (here = 100)
batch_size <- 100

# if our full sample is 4500
# then our no. of time steps is 45
N <- 4500
T <- N/batch_size
```

```{r}
# pi is a vector of probabilities
# size is the number of observations
# C is number of arms
assign_treatment <- function(pi, profile_prob, size, C=8) {
  df_fake <- data.frame(matrix(0, ncol = C, nrow = (size)))
  names(df_fake) <- str_c("Q", 1:C)
  df_fake["profile"] <- 0

  # want to get even distribution within each batch
  for (i in 1:batch_size) {
    rnum <- runif(1)
    profile <- case_when(
        rnum < pi[1] ~ 1,
        rnum >= pi[1] & rnum < sum(pi[1:2]) ~ 2,
        rnum >= sum(pi[1:2]) & rnum < sum(pi[1:3]) ~ 3,
        rnum >= sum(pi[1:3]) & rnum < sum(pi[1:4]) ~ 4,
        rnum >= sum(pi[1:4]) & rnum < sum(pi[1:5]) ~ 5,
        rnum >= sum(pi[1:5]) & rnum < sum(pi[1:6]) ~ 6,
        rnum >= sum(pi[1:6]) & rnum < sum(pi[1:7]) ~ 7,
        rnum >= sum(pi[1:7]) ~ 8
      )
    # alternative method
    #profile <- match(1, rmultinom(1, size = 1, prob = pi))
    
    # this is to simulate fake data
    df_fake[i, str_c("Q", profile)] <- rbinom(1, 1, profile_prob[profile])
    df_fake[i, "profile"] <- profile
  }
  return(df_fake)
}
```

```{r}
profile_prob <- c(0.9, 0.5, 0.3, 0.5, 0.4, 0.7, 0.5, 0.44)
pi <- rep(0.125, 8)
pc_survey_fake <- assign_treatment(pi, profile_prob, batch_size)
pc_survey_fake %>% 
  head()
```

### Examine fake data

* Check distribution of profiles overll and by batch

```{r}
pc_survey_fake %>%
  ggplot(aes(profile)) +
  geom_histogram()  +
  theme_classic()

pc_survey_fake %>%
  pivot_longer(-c(profile)) %>%
  group_by(profile) %>%
  summarize(mean = mean(value)) %>%
  ggplot(aes(profile, mean)) +
  geom_col()  +
  theme_classic()

pc_survey_fake %>%
  # share of respondents who select
  # the younger candidate
  pivot_longer(-c(profile)) %>%
  group_by(profile) %>%
  summarize(
    mean = mean(value),
    total = sum(value)
  )
```

## Normal Experiment

```{r}
# init parameters
c <- 8
profiles_selected <- integer(0)
numbers_of_rewards_1 <- integer(c) # vector for each arm
numbers_of_rewards_0 <- integer(c)
total_reward <- 0
pi <- lst(rep(0.125, 8)) # init uniform treatment assignment
R <- 1000
regret <- lst()

df_list <- tibble()
for (n in 1:T) { # for T runs (in our case 45)
  # subset data into 100 obs and assign treatment randomly
  df_subset <- assign_treatment(pi[[n]], profile_prob, batch_size, C=c)
  df_list <- bind_rows(df_list, df_subset)
  stopifnot(nrow(df_subset) == batch_size)
  max_random <- 0
  profile_val <- 1
  
  # iterate through all context arms
  for (i in 1:c) {
    # draw from prior distribution for theta
    # for each arm
    # alpha, beta parameters depend
    # on the number of prior rewards
    random_theta <- rbeta(
      n = 1,
      shape1 = numbers_of_rewards_1[i] + 1,
      shape2 = numbers_of_rewards_0[i] + 1
    )

    if (random_theta > max_random) {
      # if arm is better than all the previous ones
      # this arm is selected, choose theta
      max_random <- random_theta
      profile_val <- i
    }
  }

  # after selecting the argmax, add to selected profile
  profiles_selected <- append(profiles_selected, profile_val)

  # checking with data (what is *observed*)
  # update based on reward
  obs_outcome <- df_subset %>%
    filter(profile == profile_val) %>%
    magrittr::extract2(str_c("Q", profile_val))

  reward0 <- sum(obs_outcome == 0)
  reward1 <- sum(obs_outcome == 1)
  numbers_of_rewards_0[profile_val] <- numbers_of_rewards_0[profile_val] + reward0
  numbers_of_rewards_1[profile_val] <- numbers_of_rewards_1[profile_val] + reward1

  total_reward <- total_reward + reward1
  
  # need to generate new pi
  # will preserve as unfirom
  pi[n + 1] <- lst(c(0.125, c))
  
  # calculate per-period regret
  # expectation of best arm vs. expectation of chosen arm
  regret[n] <- max(profile_prob) - profile_prob[profile_val]
}
```

```{r}
# store impt values for comparison
pi_exp <- pi
profiles_selected_exp <- profiles_selected
regret_exp <- regret
```

### Normal experiment: plotting results

```{r, message = F, warning=F}
plot_selected_profiles(profiles_selected_exp)

plot_observed_profiles(df_list)

plot_dist(numbers_of_rewards_1, numbers_of_rewards_0, c)

plot_regret(regret_exp)
```

## Greedy Algorithm

* use the expected values of the Beta distribution

```{r}
# init parameters
c <- 8
profiles_selected <- integer(0)
numbers_of_rewards_1 <- integer(c) # vector for each arm
numbers_of_rewards_0 <- integer(c)
total_reward <- 0
pi <- lst(rep(0.125, 8)) # init uniform treatment assignment
R <- 1000

for (n in 1:T) { # for T runs (in our case 45)
  # subset data into 100 obs and assign treatment randomly
  df_subset <- assign_treatment(pi[[n]], profile_prob, batch_size, C=c)
  stopifnot(nrow(df_subset) == batch_size)
  max_random <- 0
  profile_val <- 1

  # iterate through all context arms
  for (i in 1:c) {
    # draw from prior distribution for theta
    # for each arm
    # alpha, beta parameters depend
    # on the number of prior rewards
    alpha <- numbers_of_rewards_1[i] + 1
    beta <- numbers_of_rewards_0[i] + 1
    exp_theta <- alpha/(alpha + beta)

    if (exp_theta > max_random) {
      # if arm is better than all the previous ones
      # this arm is selected, choose theta
      max_random <- random_theta
      profile_val <- i
    }
  }

  # after selecting the argmax, add to selected profile
  profiles_selected <- append(profiles_selected, profile_val)

  # checking with data (what is *observed*)
  # update based on reward
  obs_outcome <- df_subset %>%
    filter(profile == profile_val) %>%
    magrittr::extract2(str_c("Q", profile_val))

  reward0 <- sum(obs_outcome == 0)
  reward1 <- sum(obs_outcome == 1)
  numbers_of_rewards_0[profile_val] <- numbers_of_rewards_0[profile_val] + reward0
  numbers_of_rewards_1[profile_val] <- numbers_of_rewards_1[profile_val] + reward1

  total_reward <- total_reward + reward1
  
  # need to generate new pi
  # this is the probability each arm is the best
  # we can do this by simulating R random samples and calculating the probability
  # that each arm is optimal
  
  # Approximate Thompson Sampling probabilities (i.e. prob. that each arm is maximal)
  # (this is from Offer-Westort et al.)
  # 1) Sample from each of the posteriors R times
  draws <- replicate(R, rbeta(c, numbers_of_rewards_1, numbers_of_rewards_0))
  # 2) Check how many times each arm was maximal
  argmax <- apply(draws, 2, which.max)
  # 3) Tally up the probabilities
  pi[n + 1] <- lst(table(cut(argmax, 0:c)) / R)
}
```

```{r}
# store impt values for comparison
pi_greedy <- pi
profiles_selected_greedy <- profiles_selected
```

### Greedy Bernoulli: plotting results

```{r, warning=F}
plot_selected_profiles(profiles_selected)

plot_observed_profiles(df_list)

plot_dist(numbers_of_rewards_1, numbers_of_rewards_0, c)
```

## Thompson Sampling

```{r}
# init parameters
c <- 8
profiles_selected <- integer(0)
numbers_of_rewards_1 <- integer(c) # vector for each arm
numbers_of_rewards_0 <- integer(c)
total_reward <- 0
pi <- lst(rep(0.125, 8)) # init uniform treatment assignment
R <- 1000
regret <- lst()

df_list <- tibble()
for (n in 1:T) { # for T runs (in our case 45)
  # subset data into 100 obs and assign treatment randomly
  df_subset <- assign_treatment(pi[[n]], profile_prob, batch_size, C=c)
  df_list <- bind_rows(df_list, df_subset)
  stopifnot(nrow(df_subset) == batch_size)
  max_random <- 0
  profile_val <- 1
  
  # iterate through all context arms
  for (i in 1:c) {
    # draw from prior distribution for theta
    # for each arm
    # alpha, beta parameters depend
    # on the number of prior rewards
    random_theta <- rbeta(
      n = 1,
      shape1 = numbers_of_rewards_1[i] + 1,
      shape2 = numbers_of_rewards_0[i] + 1
    )

    if (random_theta > max_random) {
      # if arm is better than all the previous ones
      # this arm is selected, choose theta
      max_random <- random_theta
      profile_val <- i
    }
  }

  # after selecting the argmax, add to selected profile
  profiles_selected <- append(profiles_selected, profile_val)

  # checking with data (what is *observed*)
  # update based on reward
  obs_outcome <- df_subset %>%
    filter(profile == profile_val) %>%
    magrittr::extract2(str_c("Q", profile_val))

  reward0 <- sum(obs_outcome == 0)
  reward1 <- sum(obs_outcome == 1)
  numbers_of_rewards_0[profile_val] <- numbers_of_rewards_0[profile_val] + reward0
  numbers_of_rewards_1[profile_val] <- numbers_of_rewards_1[profile_val] + reward1

  total_reward <- total_reward + reward1
  
  # need to generate new pi
  # this is the probability each arm is the best
  # we can do this by simulating R random samples and calculating the probability
  # that each arm is optimal
  
  # Approximate Thompson Sampling probabilities (i.e. prob. that each arm is maximal)
  # (this is from Offer-Westort et al.)
  # 1) Sample from each of the posteriors R times
  draws <- replicate(R, rbeta(c, numbers_of_rewards_1, numbers_of_rewards_0))
  # 2) Check how many times each arm was maximal
  argmax <- apply(draws, 2, which.max)
  # 3) Tally up the probabilities
  pi[n + 1] <- lst(table(cut(argmax, 0:c)) / R)
  
  # calculate per-period regret
  # expectation of best arm vs. expectation of chosen arm
  regret[n] <- max(profile_prob) - profile_prob[profile_val]
}
```

```{r}
# store impt values for comparison
pi_ts <- pi
profiles_selected_ts <- profiles_selected
regret_ts <- regret
```

### TS: plotting results

```{r, message = F, warning=F}
plot_selected_profiles(profiles_selected_ts)

plot_observed_profiles(df_list)

plot_dist(numbers_of_rewards_1, numbers_of_rewards_0, c)

plot_regret(regret_ts)
```

## Epsilon Greedy

* From Offer-Westort et al. (2020): in the Epsilon Greedy algorithm, we identify which arm has the highest mean outcome. We assign treatment to that arm in (1−\epsilon)  portion of assignments, and assign treatment uniformly at random in the remaining \epsilon assignments.
* Note: can also allow epsilon to decay over time

```{r}
# init parameters
c <- 8
eps <- 0.1
profiles_selected <- integer(0)
numbers_of_rewards_1 <- integer(c) # vector for each arm
numbers_of_rewards_0 <- integer(c)
total_reward <- 0
pi <- lst(rep(0.125, 8)) # init uniform treatment assignment
R <- 1000

for (n in 1:T) { # for T runs (in our case 45)
  # subset data into 100 obs and assign treatment randomly
  df_subset <- assign_treatment(pi[[n]], profile_prob, batch_size, C=c)
  df_list <- bind_rows(df_list, df_subset)
  stopifnot(nrow(df_subset) == batch_size)
  max_random <- 0
  profile_val <- 1

  # identify which arm has highest mean outcome
  sorted_df <- df_subset %>% 
    pivot_longer(-profile, names_to = "arm", values_to = "reward") %>% 
    group_by(profile) %>% 
    summarize(mean_reward = mean(reward)) %>% 
    arrange(desc(mean_reward))
  
  # highest arm is the first one
  profile_val <- sorted_df[1,][["profile"]]
  
  # after selecting the argmax, add to selected profile
  profiles_selected <- append(profiles_selected, profile_val)

  # checking with data (what is *observed*)
  # update based on reward
  obs_outcome <- df_subset %>%
    filter(profile == profile_val) %>%
    magrittr::extract2(str_c("Q", profile_val))

  reward0 <- sum(obs_outcome == 0)
  reward1 <- sum(obs_outcome == 1)
  numbers_of_rewards_0[profile_val] <- numbers_of_rewards_0[profile_val] + reward0
  numbers_of_rewards_1[profile_val] <- numbers_of_rewards_1[profile_val] + reward1

  total_reward <- total_reward + reward1
  
  # need to generate new pi
  pi_n <- rep(eps, c)
  pi_n[profile_val] <- 1 - eps
  pi[n+1] <- lst(pi_n)
}
```

```{r}
# store impt values for comparison
pi_eg <- pi
profiles_selected_eg <- profiles_selected
```

### Epsilon-Greedy: plotting results

```{r, message = F, warning=F}
plot_selected_profiles(profiles_selected_eg, name="Epsilon Greedy")

plot_observed_profiles(df_list)

plot_dist(numbers_of_rewards_1, numbers_of_rewards_0, c)
```

## UCB (Upper Confidence Bound)

* observe sample means under each of the arms, and then we compute uncertainty bounds around each of those estimates (note that these are not the same as confidence intervals under a normal approximation)
* goal is to reduce uncertainty

```{r}
# init parameters
c <- 8
profiles_selected <- integer(0)
numbers_of_rewards_1 <- integer(c) # vector for each arm
numbers_of_rewards_0 <- integer(c)
total_reward <- 0
pi <- lst(rep(0.125, 8)) # init uniform treatment assignment
R <- 1000

for (n in 1:T) { # for T runs (in our case 45)
  # subset data into 100 obs and assign treatment randomly
  df_subset <- assign_treatment(pi[[n]], profile_prob, batch_size, C=c)
  df_list <- bind_rows(df_list, df_subset)
  stopifnot(nrow(df_subset) == batch_size)
  max_random <- 0
  profile_val <- 1

  # identify ucb arm
  sorted_df <- df_subset %>% 
    pivot_longer(-profile, names_to = "arm", values_to = "success") %>% 
    group_by(profile) %>% 
    summarize(successes = sum(success),
              trials = n()) %>% 
    ungroup() %>% 
    mutate(ucb_arm = successes / trials + sqrt(1 * log(sum(trials)) / trials)) %>% 
    arrange(desc(ucb_arm))
  
  # update selected arm
  profile_val <- sorted_df[1,][["profile"]]
  
  # after selecting the argmax, add to selected profile
  profiles_selected <- append(profiles_selected, profile_val)

  # checking with data (what is *observed*)
  # update based on reward
  obs_outcome <- df_subset %>%
    filter(profile == profile_val) %>%
    magrittr::extract2(str_c("Q", profile_val))

  reward0 <- sum(obs_outcome == 0)
  reward1 <- sum(obs_outcome == 1)
  numbers_of_rewards_0[profile_val] <- numbers_of_rewards_0[profile_val] + reward0
  numbers_of_rewards_1[profile_val] <- numbers_of_rewards_1[profile_val] + reward1

  total_reward <- total_reward + reward1
  
  # need to generate new pi
  pi_n <- rep(0, c)
  pi_n[profile_val] <- 1
  pi[n+1] <- lst(pi_n)
}
```

```{r}
# store impt values for comparison
pi_ucb <- pi
profiles_selected_ucb <- profiles_selected
```

### Epsilon-Greedy: plotting results

```{r, message = F, warning=F}
plot_selected_profiles(profiles_selected_ucb, name="UCB")

plot_observed_profiles(df_list)

plot_dist(numbers_of_rewards_1, numbers_of_rewards_0, c)
```

## Compare treatment assignments over time

```{r}
tibble('pi' = map(2:T, ~pi_greedy[[.]]),
       'arm' = map(2:T, ~c(1:8))) %>% 
  mutate(id = row_number()) %>% 
  unnest(cols = c(pi, arm)) %>% 
  ggplot(aes(id, pi)) +
  geom_line() +
  facet_wrap(~arm) +
  theme_classic() +
  labs(title = "Greedy Bernoulli Treatment Assignments over time",
       x = "Time",
       y = "Probability of being the best arm")

tibble('pi' = map(2:T, ~pi_ts[[.]]),
       'arm' = map(2:T, ~c(1:8))) %>% 
  mutate(id = row_number()) %>% 
  unnest(cols = c(pi, arm)) %>% 
  ggplot(aes(id, pi)) +
  geom_line() +
  facet_wrap(~arm) +
  theme_classic() +
  labs(title = "TS Treatment Assignments over time",
       x = "Time",
       y = "Probability of being the best arm")

tibble('pi' = map(2:T, ~pi_eg[[.]]),
       'arm' = map(2:T, ~c(1:8))) %>% 
  mutate(id = row_number()) %>% 
  unnest(cols = c(pi, arm)) %>% 
  ggplot(aes(id, pi)) +
  geom_line() +
  facet_wrap(~arm) +
  theme_classic() +
  labs(title = "Epsilon-Greedy Treatment Assignments over time",
       x = "Time",
       y = "Probability of being the best arm")
```

