---
title: "Process Qualtrics Data and Treatment Assignment Updating for Political Candidates Survey"
output: rmarkdown::github_document
date: "2023-11-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup Data Using Qualtrics API

* Load data directly from Qualtrics
* Can store API Key and credentials in `.renviron`

```{r, load packages, message=F, warning=F}
library(tidyverse)
library(qualtRics)
library(magrittr)
library(assertr)
library(RColorBrewer)

knitr::opts_chunk$set(cache.extra = 2023)

source("./_functions/data_cleaning.R")
source("./_functions/experiment_functions.R")
source("./_functions/plotting_functions.R")
source("./_functions/populate_probs.R")
```

```{r, setup qualtrics}
url <- str_glue("https://{config$datacenter_id}.qualtrics.com")
survey_name <- "Political Candidates"
survey_id <- config$pol_candidates_survey_id

# can comment out after running once
# qualtrics_api_credentials(api_key = config$api_key,
#                           base_url = url,
#                           install = TRUE,
#                           overwrite=T)

df <- load_qualtrics(survey_name)
df %>%
  head()
```

## Survey Validation

```{r}
df %>%
  distinct(Status)

# filter for test data
# replace with real data when running hte survey
df <- df %>%
  filter_test_data()

# check consent means their responses are missing
df %>%
  check_consent()

# check that all completed
df %>%
  check_completion()

df %>%
  # note: question was different prior to 11/06
  filter(StartDate >= lubridate::ymd("2023-11-06")) %>%
  check_location_screen()
```

```{r}
# create profile variable
df <- df %>%
  create_profile_var()

df %>%
  ggplot(aes(profile)) +
  geom_histogram() +
  theme_classic()
```

## Clean Qualtrics Data
```{r}
df_clean <- clean_qualtrics_data(df)

# each question number is the random ordering of the context attributes
df_clean %>%
  select(candidate_response, str_c("Q", 1:8), profile) %>%
  verify(!is.na(candidate_response))
```

## Clean Data Validation

```{r}
# check every respondent has exactly one non-missing value
df_clean %>%
  select(str_c("Q", 1:8)) %>%
  mutate(across(everything(), ~ as.character(.))) %>%
  is.na() %>%
  rowSums(na.rm = T)

# check row total is not more than 1
# 0 if selected the older candidate
df_clean %>%
  select(str_c("Q", 1:8)) %>%
  rowSums(na.rm = T) %>%
  is_weakly_less_than(1) %>%
  all() %>%
  stopifnot()
```


# Update treatment probabilities
* TODO: update this so it generates a vector of probabilities based on **real** data
* TODO: automate probabilities in Qualtrics
* TODO: remove fake data/simulated data and keep only in `run_simulated_qulatrics.Rmd`

```{r, global params}
# define with comments
num_profiles <- 8 # number of profiles
num_sim <- 1000 # number of simulations for Monte Carlo simulation
eps <- 0.1 # for epsilon greedy alg

N <- 4500 # total number of observations
num_batches <- N / batch_size # total number of batches
pi_init <- cumsum(rep(0.125, 8)) # initial pi cdf
```

```{r, message=F, warning=T}
# generate pi on real data
output <- update_ts(df_clean, num_sim, num_profiles, num_outcome1, num_outcome0, cdf = TRUE)
output

# vector of new probabilities
# assign names (impt for checking replacement in survey flow)
pi_ts <- pi_init[1:(num_profiles-1)] 
pi_ts <- output$pi[1:(num_profiles-1)]
names(pi_ts) <- str_c("pi", 1:(num_profiles-1))
```
```{r}
# Retrieve current survey flow (full structure)
current_flow <- get_survey_flow(config$api_token, survey_id, config$datacenter_id)
# Extract just the flow part for modification
current_flow_data <- current_flow$result$Flow

# Update the flow data with new probabilities
modified_flow_data <- update_flow_with_probabilities(
  current_flow_data,
  "FL_10",
  pi_ts,
  names(pi_ts)
)

# Reconstruct the full survey configuration with the modified flow part
current_flow$result$Flow <- modified_flow_data

# Convert the entire modified survey configuration to JSON
json_payload <- toJSON(current_flow$result, auto_unbox = TRUE)

# Make the PUT request to update the survey flow
update_response <- update_survey_flow(config$api_token, survey_id, current_flow$result, config$datacenter_id)

# Check the response
print(update_response)
```




