---
title: ""
format: 
  html:
    embed-resources: true
    page-layout: full
  pdf: default
format-links: false
---

```{r}
#| echo: false
#| message: false
#| warning: false
library(dplyr)
library(ggplot2)
library(gtsummary)
library(here)
library(modelr)
library(readr)
library(tidyr)
```

# Summary statistics

**Note:** Responses that failed our quality checks have been filtered out which
is why the **N**s total to somewhat less than our full sample of 10,000.

```{r}
#| echo: false
#| message: false
#| warning: false

clean_ethnicity <- function(x) {
  x |>
    mutate(
      ethnicity_cleaned = factor(
        ethnicity,
        levels = c("hisp_latin_spanish_no", "hisp_latin_spanish_yes"),
        labels = c(
          "Not Hispanic or Latino",
          "Hispanic or Latino"
        )
      )
    )
}

clean_race <- function(x) {
  levs <- c(
    "race_white",
    "race_black",
    "race_asian",
    "race_aian",
    "race_nhpi",
    "race_other",
    "race_multiracial"
  )
  labs <- c(
    "White",
    "Black or African American",
    "Asian",
    "American Indian or Alaska Native",
    "Native Hawaiian or Pacific Islander",
    "Other",
    "Two or More Races"
  )
  x |>
    mutate(
      race_cleaned = case_when(
        is.na(race) ~ NA,
        race %in% levs ~ race,
        TRUE ~ "race_multiracial"
      ),
      race_cleaned = factor(race_cleaned, levels = levs, labels = labs)
    )
}

clean_sex <- function(x) {
  x |>
    mutate(
      sex_cleaned = factor(
        sex,
        levels = c("female", "male"),
        labels = c("Female", "Male")
      )
    )
}

clean_responses <- function(x) {
  x |>
    clean_ethnicity() |>
    clean_race() |>
    clean_sex()
}

# Import data
responses <- read_csv(here("data/immigrants_main_response.csv"))
responses_validation_max <- read_csv(here("data/immigrants_max_response.csv"))
responses_validation_min <- read_csv(here("data/immigrants_min_response.csv"))

# Clean data
responses <- responses |>
  clean_responses() |>
  mutate(phase_cleaned = "Main")

responses_validation_max <- responses_validation_max |>
  clean_responses() |>
  mutate(phase = "Max", phase_cleaned = phase)

responses_validation_min <- responses_validation_min |>
  clean_responses() |>
  mutate(phase = "Min", phase_cleaned = phase)
```

```{r}
#| echo: false
#| message: false
#| warning: false

responses |>
  bind_rows(responses_validation_max) |>
  bind_rows(responses_validation_min) |>
  filter(!garbage) |>
  select(phase_cleaned, age, ends_with("_cleaned")) |>
  tbl_summary(
    by = phase_cleaned,
    label = list(
      age ~ "Age",
      ethnicity_cleaned ~ "Ethnicity",
      race_cleaned ~ "Race",
      sex_cleaned ~ "Sex"
    )
  ) |>
  bold_labels()
```

# Point Estimates (adaptive phase vs. validation phase)

**Note:** The point estimates and intervals from the main phase are means and 
95% Bayesian credible intervals estimated from the arms' posterior 
distributions. The point estimates and intervals from the validation phase are 
sample means and 95% confidence intervals calculated using t-distribution
approximation.

## Visual

```{r}
#| echo: false
#| message: false
#| warning: false

# load poststratification weights from ipums acs survey
wgts <- readRDS(here("data/ipums_strata_sizes.RDS")) |>
  mutate(
    race = as.character(race),
    race = case_when(
      race == "Native Hawaiian or Other Pacific Islander" ~
        "Native Hawaiian or Pacific Islander",
      TRUE ~ race
    ),
    race = factor(race, levels = levels(responses$race_cleaned))
  )

poststratified_estimates <- function(df, arm = "max") {
  ###############################################
  # (2) Poststratified Point Estimate
  ###############################################

  # First, drop people who choose not to disclose for any of the demographic features
  # Then align variable names with IPUMS
  df_filter <- df |>
    filter(if_all(.cols = c(race, ethnicity, sex, age), .fns = \(x) !is.na(x))) |>
    mutate(
      race = factor(as.character(race_cleaned), levels = levels(wgts$race)),
      female = sex == "female",
      hispanic = ethnicity == "hisp_latin_spanish_yes"
    )

  # exclude race categories that aren't in the data we have
  wgts <- wgts |>
    filter(race %in% unique(df_filter$race))

  # running logistic regression for predicting probability of choosing younger candidate
  glm_drop_cons_factors <- function(df, vars) {
    # need to drop factors that don't vary (e.g., only one racial category appears)
    if (length(unique(df$race)) <= 1) vars <- vars[vars != "race"]
    form <- paste0(vars, collapse = " + ")

    glm(
      formula = paste0("discriminated ~ ", form),
      family = "binomial",
      data = df
    )
  }

  # compute weighted probability of choosing the younger candidate
  # exclude race categories that aren't in the data when doing the prediction
  compute_weighted_prob <- function(df, wgts, lm) {
    # exclude race categories that aren't in the data we have
    wgts <- wgts |>
      filter(race %in% unique(as_tibble(df)[["race"]]))

    # predict probabilities
    prob <- predict(lm, newdata = wgts, type = "response")
    # return weighted mean
    return(weighted.mean(prob, w = wgts$weight))
  }

  df_models <- df_filter |>
    arrange(arm_id) |>
    group_by(arm_id) |>
    nest() |>
    mutate(glm = lapply(
      data,
      \(x) {
        glm_drop_cons_factors(
          x,
          c("race", "female", "hispanic", "age")
        )
      }
    ))

  # predict based on population weight categories
  # and then compute the weighted mean
  w_mean <- lapply(
    1:length(df_models$data),
    \(index) {
      data <- df_models[["data"]][[index]]
      model <- df_models[["glm"]][[index]]
      compute_weighted_prob(data, wgts, model)
    }
  ) |>
    unlist()

  df_post <- tibble(
    context = df_models$arm_id,
    estimate = w_mean
  )

  ###############################################
  # (3) CI with Bootstrapping
  ###############################################

  produce_bootstrap_estimates <- function(df, c_val, iter = 1000) {
    df_bootstrap <- df |>
      filter(arm_id == c_val) |>
      bootstrap(iter) |>
      mutate(glm = lapply(
        strap,
        \(x) {
          glm_drop_cons_factors(
            x,
            c("female", "hispanic", "age", "race")
          )
        }
      ))

    # produce bootstrap estimate
    bootstrap_est <- lapply(
      1:length(df_bootstrap$strap),
      \(index) {
        strap <- df_bootstrap$strap[[index]]
        model <- df_bootstrap$glm[[index]]
        compute_weighted_prob(strap, wgts, model)
      }
    ) |>
      unlist()

    # produce mean and sd
    tibble(
      se = sd(bootstrap_est),
      context = c_val
    ) |>
      select(context, se)
  }

  c_val <- unique(df_filter$arm_id)
  df_post_bootstrap <- lapply(c_val, \(x) produce_bootstrap_estimates(df_filter, c_val)) |>
    bind_rows() |>
    full_join(df_post, ., by = c("context")) |>
    # adding bootstrap standard error to calculation of ci with point estimate of the mean
    mutate(
      phase = "Poststratified",
      arm_id = case_when(.env$arm == "max" ~ 7, TRUE ~ 10),
      mu = estimate,
      lb = estimate - (qt(.975, df = nrow(df_filter) - 1) * se),
      ub = estimate + (qt(.975, df = nrow(df_filter) - 1) * se)
    ) |>
    select(arm_id, mu, lb, ub, phase)

  return(df_post_bootstrap)
}

ps_max <- poststratified_estimates(responses_validation_max, "max")
ps_min <- poststratified_estimates(responses_validation_min, "min")
ps <- bind_rows(ps_max, ps_min)
```

```{r}
#| echo: false
#| message: false
#| warning: false

# Import parameters estimated in the adaptive phase
params_main <- read_csv(here("data/immigrants_main_parameters.csv")) |>
  filter(batch_id == max(batch_id)) |>
  arrange(desc(arm_id)) |>
  filter(arm_id %in% c(10, 7)) |>
  group_by(arm_id) |>
  mutate(
    mu = alpha/(alpha + beta),
    lb = qbeta(0.025, alpha, beta),
    ub = qbeta(0.975, alpha, beta),
    phase = "Main"
  ) |>
  ungroup() |>
  select(arm_id, mu, lb, ub, phase)

# Import parameters estimated in the validation phase
params_validation <- responses_validation_max |> 
  mutate(arm_id = 7) |>
  bind_rows(
    mutate(responses_validation_min, arm_id = 10)
  ) |>
  filter(!garbage) |> 
  group_by(arm_id) |>
  summarise(
    mu = mean(discriminated),
    sd = sd(discriminated), root_n = sqrt(n()),
    df = n() - 1
  ) |> 
  ungroup() |>
  group_by(arm_id) |>
  mutate(
    lb = mu - qt(0.975, df = df)*sd/root_n,
    ub = mu - qt(0.025, df = df)*sd/root_n,
    phase = "Validation"
  ) |>
  ungroup() |>
  select(arm_id, mu, lb, ub, phase)

params_comparison <- bind_rows(params_main, params_validation) |>
  bind_rows(ps) |>
  arrange(arm_id) |>
  mutate(
    phase = factor(
      phase,
      levels = c("Main", "Validation", "Poststratified"),
      ordered = TRUE
    )
  )
```

```{r}
#| echo: false
#| fig-align: center
#| message: false
#| warning: false

ggplot(
    params_comparison,
    aes(x = factor(arm_id), y = mu, ymin = lb, ymax = ub, color = phase)
  ) +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_errorbar(width = 0.1, position = position_dodge(width = 0.3)) +
  theme_minimal() +
  labs(x = "Arm", y = expression(mu ~ "(" %+-% " 95% CI)"), color = "")
```

## Tabular

```{r}
#| echo: false
#| layout-align: center
#| message: false
#| warning: false

params_comparison |>
  mutate(
    context = case_when(arm_id == 7 ~ "Max", TRUE ~ "Min"),
    across(where(is.numeric), \(x) round(x, 3))
  ) |>
    select(context, lb, everything(), -arm_id) |>
    gt::gt() |>
    gt::cols_label(
      context = gt::md("**Context**"),
      mu = gt::md("**Mu**"),
      lb = gt::md("**Lower Bound**"),
      ub = gt::md("**Upper Bound**"),
      phase = gt::md("**Phase**")
    )
```
