---
title: ""
format: 
  html:
    embed-resources: true
    page-layout: full
  pdf: default
format-links: false
---

```{r}
#| echo: false
#| message: false
#| warning: false
library(dplyr)
library(ggplot2)
library(gtsummary)
library(here)
library(readr)
```

# Summary statistics

**Note:** Responses that failed our quality checks have been filtered out which
is why the **N**s total to somewhat less than our full sample of 10,000.

```{r}
#| echo: false
#| message: false
#| warning: false

clean_ethnicity <- function(x) {
  x |>
    mutate(
      ethnicity_cleaned = factor(
        ethnicity,
        levels = c("hisp_latin_spanish_no", "hisp_latin_spanish_yes"),
        labels = c(
          "Hispanic or Latino",
          "Not Hispanic or Latino"
        )
      )
    )
}

clean_race <- function(x) {
  levs <- c(
    "race_white",
    "race_black",
    "race_asian",
    "race_aian",
    "race_nhpi",
    "race_other",
    "race_multiracial"
  )
  labs <- c(
    "White",
    "Black or African American",
    "Asian",
    "American Indian or Alaskan Native",
    "Native Hawaiian or Pacific Islander",
    "Other",
    "Multiracial"
  )
  x |>
    mutate(
      race_cleaned = case_when(
        is.na(race) ~ NA,
        race %in% levs ~ race,
        TRUE ~ "race_multiracial"
      ),
      race_cleaned = factor(race_cleaned, levels = levs, labels = labs)
    )
}

clean_sex <- function(x) {
  x |>
    mutate(
      sex_cleaned = factor(
        sex,
        levels = c("female", "male"),
        labels = c("Female", "Male")
      )
    )
}

clean_responses <- function(x) {
  x |>
    clean_ethnicity() |>
    clean_race() |>
    clean_sex()
}

# Import data
responses <- read_csv(here("data/immigrants_main_response.csv"))
responses_validation_max <- read_csv(here("data/immigrants_max_response.csv"))
responses_validation_min <- read_csv(here("data/immigrants_min_response.csv"))

# Clean data
responses <- responses |>
  clean_responses() |>
  mutate(phase_cleaned = "Main")

responses_validation_max <- responses_validation_max |>
  clean_responses() |>
  mutate(phase = "Max", phase_cleaned = phase)

responses_validation_min <- responses_validation_min |>
  clean_responses() |>
  mutate(phase = "Min", phase_cleaned = phase)
```

```{r}
#| echo: false
#| message: false
#| warning: false

responses |>
  bind_rows(responses_validation_max) |>
  bind_rows(responses_validation_min) |>
  filter(!garbage) |>
  select(phase_cleaned, ends_with("_cleaned")) |>
  tbl_summary(
    by = phase_cleaned,
    label = list(
      ethnicity_cleaned ~ "Ethnicity",
      race_cleaned ~ "Race",
      sex_cleaned ~ "Sex"
    )
  ) |>
  bold_labels()
```

# Point Estimates (adaptive phase vs. validation phase)

**Note:** The point estimates and intervals from the main phase are means and 
95% Bayesian credible intervals estimated from the arms' posterior 
distributions. The point estimates and intervals from the validation phase are 
sample means and 95% confidence intervals calculated using t-distribution
approximation.

```{r}
#| echo: false
#| message: false
#| warning: false

params_main <- read_csv(here("data/immigrants_main_parameters.csv")) |>
  filter(batch_id == max(batch_id)) |>
  arrange(desc(arm_id)) |>
  filter(arm_id %in% c(10, 7)) |>
  group_by(arm_id) |>
  mutate(
    mu = alpha/(alpha + beta),
    lb = qbeta(0.025, alpha, beta),
    ub = qbeta(0.975, alpha, beta),
    phase = "Main"
  ) |>
  ungroup() |>
  select(arm_id, mu, lb, ub, phase)

params_validation <- responses_validation_max |> 
  mutate(arm_id = 7) |>
  bind_rows(
    mutate(responses_validation_min, arm_id = 10)
  ) |>
  filter(!garbage) |> 
  group_by(arm_id) |>
  summarise(
    mu = mean(discriminated),
    sd = sd(discriminated), root_n = sqrt(n()),
    df = n() - 1
  ) |> 
  ungroup() |>
  group_by(arm_id) |>
  mutate(
    lb = mu - qt(0.975, df = df)*sd/root_n,
    ub = mu - qt(0.025, df = df)*sd/root_n,
    phase = "Validation"
  ) |>
  ungroup() |>
  select(arm_id, mu, lb, ub, phase)

params_comparison <- bind_rows(params_main, params_validation) |> arrange(arm_id)
```

```{r}
#| echo: false
#| layout-align: center
#| message: false
#| warning: false

ggplot(
    params_comparison,
    aes(x = factor(arm_id), y = mu, ymin = lb, ymax = ub, color = phase)
  ) +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_errorbar(width = 0.1, position = position_dodge(width = 0.3)) +
  theme_minimal() +
  labs(x = "Arm", y = expression(mu ~ "(" %+-% " 95% CI)"), color = "")
```

