---
title: "Simulation for Adaptive Conjoint Appendix"
format: pdf
---

```{r, message = F, warning = F}
library(tidyverse)
theme_set(theme_bw())
library(foreach)
library(doParallel)
library(doRNG)
cl <- makeCluster(detectCores())
registerDoParallel(cl)
```

Define true parameter to learn.

```{r}
theta <- tibble(
  arm = 1:5,
  theta = seq(.3,.7,.1)
)
```

Define an estimator that samples arms with equal probabilities that are counterbalanced to have equal sample sizes.

```{r}
equal_probability_estimator <- function(total_n, arms = theta$arm) {
  if (length(arms) == 1) {
    data_aggregated <- tibble(id = 1:total_n) |>
      mutate(
        arm = arms
      )
  } else {
    data_aggregated <- tibble(id = 1:total_n) |>
      mutate(
        arm = replicate(
          n = ceiling(total_n / length(arms)),
          expr = sample(arms)
        )[1:total_n]
      )
  }
  data_aggregated <- data_aggregated |>
    # Merge in true choice probabilities
    left_join(theta, by = join_by(arm)) |>
    # Simulate the choice outcome
    mutate(y = rbinom(n(), 1, theta)) |>
    # Count choice outcomes by arm
    group_by(arm) |>
    summarize(y1 = sum(y == 1), y0 = sum(y == 0))
  
  # Simulate from posterior beta distribution
  estimates <- foreach(sim_rep = 1:1e3, .combine = "rbind") %do% {
    data_aggregated |>
      mutate(
        # Simulate theta from the posterior
        theta = rbeta(n(), 1 + y1, 1 + y0)
      ) |>
      # Determine which theta is biggest in this sim_rep
      arrange(theta) |>
      mutate(is_extreme = 1:n() == n())
  } |>
    select(arm, theta, is_extreme) |>
    group_by(arm) |>
    summarize(
      theta_hat = mean(theta),
      theta_ci_min = quantile(theta, .025),
      theta_ci_max = quantile(theta, .975),
      p_max = mean(is_extreme)
    )
  
  return(estimates)
}
```

Define an adaptive estimator with `warmup_n` cases used for warmup and `adaptive_n` cases used for adaptive estimation.

```{r}
adaptive_estimator <- function(warmup_n, adaptive_n, arms = theta$arm) {
  
  # Warm-up phase
  data_aggregated <- tibble(id = 1:warmup_n) |>
    # Assign arms systematically with equal probabilities
    mutate(
      arm = replicate(
        n = ceiling(warmup_n / length(arms)),
        expr = sample(arms)
      )[1:warmup_n]
    ) |>
    # Merge in true parameter values
    left_join(theta, by = join_by(arm)) |>
    # Simulate data
    mutate(y = rbinom(n(), 1, theta)) |>
    group_by(arm) |>
    summarize(y1 = sum(y), y0 = sum(1 - y)) |>
    select(arm, y1, y0)
  
  for (i in 1:adaptive_n) {
    # Adaptively update 1 arm with 1 new data point
    updated_arm <- data_aggregated |>
      # Simulate 1 posterior theta value for each arm
      mutate(sim_theta = rbeta(n(), 1 + y1, 1 + y0)) |>
      # Keep the largest theta value
      arrange(sim_theta) |>
      slice_tail(n = 1) |>
      # Draw data from that arm, using true (unknown) theta
      left_join(theta, by = join_by(arm)) |>
      mutate(y = rbinom(1, 1, theta)) |>
      # Update aggregate data
      mutate(
        y1 = y1 + y, 
        y0 = y0 + (1 - y)
      ) |>
      select(arm, y1, y0)
    
    # Update the aggregated data for that arm
    data_aggregated <- data_aggregated |>
      filter(arm != updated_arm$arm) |>
      bind_rows(updated_arm)
  }
  
  # Simulate from posterior beta distribution
  estimates <- foreach(sim_rep = 1:1e3, .combine = "rbind") %do% {
    data_aggregated |>
      mutate(
        # Simulate a value of theta from the posterior
        theta = rbeta(n(), 1 + y1, 1 + y0)
      )  |>
      # Determine which theta is biggest in this sim_rep
      arrange(theta) |>
      mutate(is_extreme = 1:n() == n())
  } |>
    select(arm, theta, is_extreme) |>
    group_by(arm) |>
    summarize(
      theta_hat = mean(theta),
      theta_ci_min = quantile(theta, .025),
      theta_ci_max = quantile(theta, .975),
      p_max = mean(is_extreme)
    )
  
  return(estimates)
}
```


Write a function that calls the above to make all 2 $\times$ 2 cells of estimates that involve:

1) One-sample estimation vs
2) Selection followed by a separate validation set

and

1) Equal probability randomization for selection
2) Warmup + adaptive randomization for selection

Throughout, the total sample size is kept constant. As a result, an estimator that does not involve a selection step gets to use more observations for the single-sample estimation.

```{r}
simulate_selection_bias <- function(total_n, selection_n, warmup_n) {
  
  # Define sample sizes implicit in arguments
  validation_n <- total_n - selection_n
  adaptive_n <- selection_n - warmup_n
  
  # Carry out equal and adaptive estimation in the full sample (no split)
  equal_full_sample <- equal_probability_estimator(total_n = total_n)
  adaptive_full_sample <- adaptive_estimator(
    warmup_n = warmup_n, 
    adaptive_n = total_n - warmup_n, 
    arms = theta$arm
  )
  
  # Carry out selection estimation in a selection sample
  equal_selection_sample <- equal_probability_estimator(total_n = selection_n)
  adaptive_selection_sample <- adaptive_estimator(
    warmup_n = warmup_n, 
    adaptive_n = adaptive_n, 
    arms = theta$arm
  )
  
  # Determine which arm had the max probability of being the max in each estimator
  equal_selected <- equal_selection_sample |> arrange(p_max) |> slice_tail(n = 1) |> pull(arm)
  adaptive_selected <- adaptive_selection_sample |> arrange(p_max) |> slice_tail(n = 1) |> pull(arm)
  
  # Carry out validation estimation for selected cases
  equal_validation <- equal_probability_estimator(
    total_n = validation_n, 
    arms = equal_selected
  )
  adaptive_validation <- equal_probability_estimator(
    total_n = validation_n, 
    arms = adaptive_selected
  )
  
  return(
    equal_full_sample |> mutate(sample = "full", randomization = "equal")|>
      bind_rows(
        adaptive_full_sample |> mutate(sample = "full", randomization = "adaptive")
      ) |>
      bind_rows(
        equal_selection_sample |> mutate(sample = "selection", randomization = "equal")
      ) |>
      bind_rows(
        adaptive_selection_sample |> mutate(sample = "selection", randomization = "adaptive")
      ) |>
      bind_rows(
        equal_validation |> mutate(sample = "validation", randomization = "equal")
      )  |>
      bind_rows(
        adaptive_validation |> mutate(sample = "validation", randomization = "adaptive")
      )
  )
}
```

The code below applies the functions to carry out the simulation.

```{r}
t0 <- Sys.time()
selection_simulations <- foreach(
  rep = 1:1e3, 
  .combine = "rbind",
  .packages = c("tidyverse","foreach")
) %dorng% {
  simulate_selection_bias(
    total_n = 500, selection_n = 250, warmup_n = 50
    # bottom is slightly better for showing bias of selection set,
    # but main claims get noisier and less clear
    #total_n = 300, selection_n = 200, warmup_n = 50
  ) |>
    mutate(simulation = rep)
}
spent <- difftime(Sys.time(),t0)
```

Visualize adaptive vs equal randomization estimation in the selection set.

```{r}
forplot <- selection_simulations |>
  filter(arm == 5) |>
  filter(sample == "selection") |>
  left_join(theta, by = join_by(arm))

summaries <- forplot |>
  group_by(randomization) |>
  summarize(
    mean_interval_width = mean(theta_ci_max - theta_ci_min),
    rmse = sqrt(mean((theta_hat - .7) ^ 2)),
    bias = mean(theta_hat - .7)
  )

summaries |>
  pivot_longer(cols = -randomization) |>
  pivot_wider(names_from = "randomization") |>
  mutate(equal_over_adaptive = equal / adaptive)

forplot |>
  left_join(summaries, by = join_by(randomization)) |>
  mutate(
    randomization = case_when(
      randomization == "equal" ~ "Fixed Randomization",
      randomization == "adaptive" ~ "Adaptive Randomization"
    ),
    randomization = paste0(
      randomization,"\nRMSE = ",
      scales::label_number(accuracy = .001)(rmse),
      "\nMean CI Width = ",
      scales::label_number(accuracy = .001)(mean_interval_width)
    )
  ) |>
  ggplot(aes(
    y = theta_hat, 
    x = simulation, 
    ymin = theta_ci_min, 
    ymax = theta_ci_max
  )) +
  geom_errorbar(width = 0, color = "gray") +
  geom_point(size = .5) +
  geom_hline(
    yintercept = theta |> filter(arm == 5) |> pull(theta),
    linetype = "dashed"
  ) +
  facet_wrap(~randomization, nrow = 1) +
  xlab("Simulation Index") +
  ylab("Estimate for Context 5") +
  theme(panel.grid = element_blank())
```

With adaptive estimation, on average the posterior probability that Arm 5 is the correct arm is higher.

```{r}
selection_simulations |>
  filter(sample == "selection") |>
  filter(arm == 5) |>
  group_by(randomization) |>
  summarize(p_max = mean(p_max))
```

This means you choose to focus on the correct arm at a slightly higher rate at a given sample size.

```{r}
selection_estimates <- selection_simulations |>
  filter(sample == "selection") |>
  # Keep only the arm that would be selected as the max
  group_by(randomization, simulation) |>
  filter(p_max == max(p_max))
selection_estimates |>
  # Summarize how often that is correct
  group_by(randomization) |>
  summarize(correct_arm = mean(arm == 5))
```

There is a potential problem of bias induced by selecting the highest-estimate arm. This can be solved with a validation set, but at a cost to efficiency.

```{r}
bias_selection <- selection_estimates |>
  group_by(randomization) |>
  summarize(
    bias = mean(theta_hat - .7),
    bias_se = sd(theta_hat) / sqrt(n())
  ) |>
  mutate(sample = "Selection Set Estimates")
bias_validation <- selection_simulations |>
  filter(sample == "validation") |>
  group_by(randomization) |>
  summarize(
    bias = mean(theta_hat - .7),
    bias_se = sd(theta_hat) / sqrt(n())
  ) |>
  mutate(sample = "Validation Set Estimates")
bias_selection |>
  bind_rows(bias_validation) |>
  ggplot(aes(
    x = randomization, 
    y = bias, 
    ymin = bias - qnorm(.975) * bias_se,
    ymax = bias + qnorm(.975) * bias_se
  )) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point() +
  geom_errorbar(width = .2) +
  facet_wrap(~sample)
```

Visualize adaptive vs equal randomization estimation in the validation set.

```{r}
forplot <- selection_simulations |>
  filter(sample == "validation")

summaries <- forplot |>
  group_by(randomization) |>
  summarize(
    mean_interval_width = mean(theta_ci_max - theta_ci_min),
    rmse = sqrt(mean((theta_hat - .7) ^ 2)),
    bias = mean(theta_hat - .7)
  )

summaries |>
  pivot_longer(cols = -randomization) |>
  pivot_wider(names_from = "randomization") |>
  mutate(equal_over_adaptive = equal / adaptive)

forplot |>
  left_join(summaries, by = join_by(randomization)) |>
  mutate(
    randomization = case_when(
      randomization == "equal" ~ "Fixed Randomization",
      randomization == "adaptive" ~ "Adaptive Randomization"
    ),
    randomization = paste0(
      randomization,"\nRMSE = ",
      scales::label_number(accuracy = .001)(rmse),
      "\nMean CI Width = ",
      scales::label_number(accuracy = .001)(mean_interval_width)
    )
  ) |>
  ggplot(aes(
    y = theta_hat, 
    x = simulation, 
    ymin = theta_ci_min, 
    ymax = theta_ci_max
  )) +
  geom_errorbar(width = 0, color = "gray") +
  geom_point(size = .5) +
  geom_hline(
    yintercept = theta |> filter(arm == 5) |> pull(theta),
    linetype = "dashed"
  ) +
  facet_wrap(~randomization, nrow = 1) +
  xlab("Simulation Index") +
  ylab("Estimated Maximum Theta") +
  theme(panel.grid = element_blank())
```

